Geometric variational problems

### 2020-11-09 NSF2020/Description/specificaim2_ver2.tex
RJR

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement of the solvent.
In Specific Aim 2, we include hydrodynamics effects by embedding the particles in solvent. 
The fluid is modeled by the Stokes equations an incompressible, zero Reynolds-number fluid, and coupled to the screened-Laplace equation \eqref{SL}
through the viscous and hydrophobic stress balance. We develope a numerical solver for the mobility problem. 
This leads to simulations of bilayers in external flows where the bilayers are represented by a self-assebled collection of rigid particles. 
Finally, we work toward three-dimensional solvers with application in the capillary driven fabrication of micro-materials.

Hydrodynamic effects are relevant detail 
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929}. 
The hydrophobic attraction still causes the particles to self-assemble, but the trajectories and even configurations are changed
by the fact that the particles must now move through an ambient environment. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

@@

The particle velocity and rotation $({\bf v}_i, {\bf w}_i)$ update the particle configurations:
\begin{alignat}{3}
\label{Hupdate1}   \frac{\dif {\bf a}_i}{\dif t} = {\bf v}_i,      \quad  \frac{\dif {\bf R}_i}{\dif t}  = A({\bf w}_i),\quad i = 1, \dots, N.
\end{alignat}
Recall that $A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
Also, ${\bf a}_i$ is the particle center the rotation matrix ${\bf R}_i$ defines the orientation.
That way, the particle configurations come from the flow maps $F_i({\bf x},t) = {\bf a}_i(t) + {\bf R_i}(t){\bf x}$ with
$P_i(t) = F_i(P_{0i},t)$ and $f_i(F_i({\bf x},t)) = f_{i0}({\bf x})$ where $P_{i0}$ and $f_{i0}$ are the $i$th reference particle
and material label respectively.

Numerically solving \eqref{St1}--\eqref{HSB2} is a nontrivial matter.
First, we have to carefully map the particle configurations because the domain $\Omega(t)$ and boundary conditions $f_i$ are constantly changing.
The particle dynamics in \eqref{Hupdate1} and in \eqref{SGD} both involve solving
screened Laplace boundary value problems on evolving domains, but differ in that \eqref{Hupdate1} has an additional Stokes flow
generating the dissipation. When there is no background flow ${\bf u}_{\infty} = {\bf 0}$, the system obeys the energy law
$\frac{\dif}{\dif t} \Phi + \int_{\Omega(t)} \mu |\nabla {\bf u} + \nabla {\bf u}^T|^2 \,dx = 0,$ 
and the particle configurations tend toward a local equilibrium.

We use \eqref{St1}--\eqref{HSB2} to simulate two-dimensional vesicle bilayer in external flows. 
The vesicle bialyer is formed by a many-body  collection of amphiphilic particles. The inner, bilayer leaflet consists
of particles pointing their hydrophobic tails away from the vesicle center; the tails of the outer leaflet point inward.
We consider two kinds of external flows: shear flow and extensional flow. 
Specifically, we deal with the background flow using the representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S(\cdot, {\bf a}_i)$ and $ R(\cdot, {\bf a}_i)$ are stokeslets and rotlets supported at the particle centers ${\bf a}_i.$
The symbol $D$ denotes the Stokes double layer integral operator (see Specific Aim 3).   
The representation \eqref{PowerMiranda} automatically satisfies \eqref{St1}, \eqref{St2}, \eqref{St4} and \eqref{HSB1}.
We solve for the unknown density $\eta : \partial \Omega \to \mathbb{R}$ by requiring that jump in viscous stress
vanished across $\partial P_i$. This jump condition guarantees that ${\bf u}$ is a rigid motion within each particle,
and by continuity, equals a rigid motion on the particle boundary as required by \eqref{St3}.

>>>>

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with elastic material properties of self-assembled collections of amphiphiles.
In Specific Aim 2, we include hydrodynamics effects by embedding the particles in a solvent. 
We divide space into the fluid phase and the rigid particle phase.
The fluid phase is modeled by the Stokes equations for an incompressible, zero Reynolds-number fluid. These are coupled to the screened-Laplace equation \eqref{SL}
through viscous and hydrophobic stress balance.
We develope a numerical solver for the mobility problem. 
This leads to simulations of bilayers in external flows where the bilayers are represented by a self-assebled collection of rigid particles. 
Finally, we work toward three-dimensional solvers with application in the capillary driven fabrication of micro-materials.

Hydrodynamic effects are a relevant detail 
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929}. 
While hydrophobic attraction causes the particles to self-assemble indepedently of the dissipation mechanism, the trajectories and
equilibrium configurations depend on the fact that the particles travel through a viscous environment. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

@@

The particle velocity and rotation $({\bf v}_i, {\bf w}_i)$ update the particle configurations:
\begin{alignat}{3}
\label{Hupdate1}   \frac{\dif {\bf a}_i}{\dif t} = {\bf v}_i,      \quad  \frac{\dif {\bf R}_i}{\dif t}  = A({\bf w}_i),\quad i = 1, \dots, N.
\end{alignat}
Recall that $A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
Also, ${\bf a}_i$ is the particle center the rotation matrix ${\bf R}_i$ defines the orientation.
That way, the particle configurations come from the flow maps $F_i({\bf x},t) = {\bf a}_i(t) + {\bf R_i}(t){\bf x}$ with
$P_i(t) = F_i(P_{0i},t)$ and $f_i(F_i({\bf x},t)) = f_{i0}({\bf x})$ where $P_{i0}$ and $f_{i0}$ are the $i$th reference particle
and material label respectively.
When there is no background flow ${\bf u}_{\infty} = {\bf 0}$, the system obeys the energy law
$\frac{\dif}{\dif t} \Phi + \int_{\Omega(t)} \mu |\nabla {\bf u} + \nabla {\bf u}^T|^2 \,dx = 0,$ 
and the particle configurations tend toward a local equilibrium.

Numerically solving \eqref{St1}--\eqref{HSB2} is a nontrivial matter.
The principle challenges are \textbf{(1)} the domain $\Omega(t)$ and boundary conditions $f_i$ are constantly changing
with changes in the particle configuration \textbf{(2)} a high degree of spatial accuracy is required to resolve the fields 
between adjacent particles and \textbf{(3)} the computational complexity grows with the number of particles. 
Specific Aim 3 describes how we address these challenges.

We simulate two-dimensional vesicle bilayers in external flows using \eqref{St1}--\eqref{HSB2}.
The vesicle bilayer is formed by a many-body collection of amphiphilic particles. There are two leaflets: 
the inner leaflet consisting of particles in contact and pointing their hydrophobic tails away from the vesicle interior;
the particles in the outer leaflet are in contact with the surrounding fluid and point their tails inward.
The particle laden flow changes the shape of the bilayer, and in return the hydrophobic stresses coming from the altered shape 
impart a force on the flow. In other words, there is a two-way coupling between the flow and bilayer shape.

We consider two kinds of external flows: shear flow and extensional flow. 
Specifically, we deal with the background flow using the representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S(\cdot, {\bf a}_i)$ and $ R(\cdot, {\bf a}_i)$ are stokeslets and rotlets supported at the particle centers ${\bf a}_i.$
The symbol $D$ denotes the Stokes double layer integral operator (see Specific Aim 3).   
The representation \eqref{PowerMiranda} automatically satisfies \eqref{St1}, \eqref{St2}, \eqref{St4} and \eqref{HSB1}.
We solve for the unknown density $\eta : \partial \Omega \to \mathbb{R}$ by requiring that jump in viscous stress
vanished across $\partial P_i$. This jump condition guarantees that ${\bf u}$ is a rigid motion within each particle,
and by continuity, equals a rigid motion on the particle boundary as required by \eqref{St3}.
